---
title: "Mediterranean Sea Fishing Activity 2019"
author: "Althea Marks"
date: "2021_5_24"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    theme: yeti
---

## Purpose

Coding test for Fisheries Analyst position at Global Fishing Watch 2021.

## Background

Data files are fishing activity in the Mediterranean Sea from 2019. Fishing hours are binned on a 0.05 degree grid. Shapefiles are of closed fishing areas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, results = "hide", warning = FALSE)
```

```{r load_libraries}
library(readr) # read in csv
library(magrittr) # pipe function
library(tidyverse) # ggplot and data wrangling
library(sp) # working with spatial data
library(rgdal) # working with vector data
library(maptools)
library(rgeos)
library(maps) # graphing maps
library(rnaturalearth) # map data source, website with descriptions
library(sf) # working with coordinates in ggplot
library(ggspatial)
library(viridisLite) # color
library(kableExtra) # table formatting
```

```{r directories_paths, warning=TRUE}
dir.create("./output") # create output directory

data_path <- file.path("data") # data path
out_path <- file.path("output") # output path
```


```{r import_data}
analyst_fishing_data <- read_csv(file = file.path(data_path, "analyst_fishing_data.csv"))
# Look at dimensions of data
dim(analyst_fishing_data)
# Look at summary stats of data
summary(analyst_fishing_data)

### Read in 2 closure shapefiles from Rproject
closure_poly_sp <- readOGR(file.path(data_path, "VME-DB_VME_GFCM_1"))
# Look at polygon
#plot(closure_poly_sp)

other_poly_sp <- readOGR(file.path(data_path, "VME-DB_VME_GFCM_4"))
# Look at polygon
#plot(other_poly_sp)
```

## Question 1: Fishing Hours

Calculate the total fishing hours, by gear type, inside both closure areas in 2019.

```{r Q1, results="markup"}
# create dataframe from shapefile for ggplot
closure_poly_df <- fortify(closure_poly_sp)
other_poly_df <- fortify(other_poly_sp)

# Graph fishing activity by fishing_class and display closure polygons
test_plot <- ggplot() +
  geom_point(data = analyst_fishing_data,
             aes(x = lon_bin, y = lat_bin,
                 color = fishing_class)) +
  geom_polygon(data = closure_poly_df,
               aes(x = long, y = lat),
               color = 'black',
               fill = NA) +
  geom_polygon(data = other_poly_df,
               aes(x = long, y = lat),
               color = "black",
               fill = NA) +
  theme_bw() +
  labs(x = "Longitude", y = "Latitude")

# test_plot

# Convert fishing dataframe to SpatialPointsDataFrame class with coordinates from columns
# Remember: Longitude listed first, then Latitude
analyst_fishing_data_sp <- analyst_fishing_data # create new object
coordinates(analyst_fishing_data_sp) <- ~ lon_bin + lat_bin # assign long & lat to coordinates
#plot(analyst_fishing_data_sp)

# Check CRS in closure & other_area shp files and fishing data
closure_poly_sp@proj4string
other_poly_sp@proj4string
analyst_fishing_data_sp@proj4string
# closure_poly_sp & other_poly_sp have same CRS, analyst_fishing_data_sp has no CRS

# Assign CRS to analyst_fishing_data_sp from closure_poly_sp
proj4string(analyst_fishing_data_sp) <- proj4string(closure_poly_sp)
# Check if same and assigned properly
analyst_fishing_data_sp@proj4string 

# Clip fishing points within closure polygon
fishing_in_closure_sp <- analyst_fishing_data_sp[closure_poly_sp,]
#plot(fishing_in_closure_sp)
fishing_in_closure <- fishing_in_closure_sp@data # extract data from sp points dataframe
# Summarize fishing hours in closure polygon
fish_hrs_closure_sum <- fishing_in_closure %>%
  group_by(fishing_class) %>% #group data by gear type
  summarise(total_fish_hrs = sum(fishing_hours)) # sum fishing hours by gear type

knitr::kable(fish_hrs_closure_sum, 
             caption = "Total Fishing Hours by Fishing Class Within 'Closure' Polygon",
             col.names = gsub("[_]", " ", names(fish_hrs_closure_sum))) %>%
  kableExtra::kable_styling(full_width = F)

# write_csv(fish_hrs_closure_sum, "./output/2019_FishHrs_Closure_ByGear.csv")

# clip fishing points within other polygon
fishing_in_other_sp <- analyst_fishing_data_sp[other_poly_sp,]
#plot(fishing_in_other_sp)
fishing_in_other <- fishing_in_other_sp@data # extract data from sp points dataframe
# summarize fishing hours in other polygon
fish_hrs_other_sum <- fishing_in_other %>% 
  group_by(fishing_class) %>% #group data by gear type
  summarise(total_fish_hrs = sum(fishing_hours)) # sum fishing hours by gear type

knitr::kable(fish_hrs_other_sum , 
             caption = "Total Fishing Hours by Fishing Class Within 'Other' Polygon",
             col.names = gsub("[_]", " ", names(fish_hrs_other_sum ))) %>%
  kableExtra::kable_styling(full_width = F)

write_csv(fish_hrs_other_sum, file.path(out_path, "2019_FishHrs_Other_ByGear.csv"))

# Combine clipped fishing data from both closed areas into single dataframe
fishing_in_both <- rbind(fishing_in_closure, fishing_in_other)

# Check what years are represented
unique(fishing_in_both$year)

# Summarize fishing within the 2 polygons ("closure", "other")
fish_hrs_both_sum <- fishing_in_both %>% 
  group_by(fishing_class) %>% # group data by gear type
  summarise(total_fish_hrs = sum(fishing_hours)) # sum fishing hours by gear type

knitr::kable(fish_hrs_both_sum, 
             caption = "Total Fishing Hours by Fishing Class Within 'Closure' and 'Other' Polygons",
             col.names = gsub("[_]", " ", names(fish_hrs_both_sum))) %>%
  kableExtra::kable_styling(full_width = F)

# Save .csv of findings to Rproject folder
# write_csv(fish_hrs_both_sum, "./output/2019_FishHrs_BothClosure_ByGear.csv")
```

## Question 2: Closure Area Maps

Provide a map or two showing the closure areas overlaid with fishing effort for trawlers and longliners.

```{r Q2_closure, results='markup'}

unique(analyst_fishing_data$fishing_class) # look at all fishing_class types
# Check with GFW, does "longliners" mean both "set_longlines" and "drifting_longlines"?

# Filter out trawler and longliner fishing data
fishing_trawl_long <- analyst_fishing_data %>% 
  filter(fishing_class %in% c("trawlers", "set_longlines", "drifting_longlines"))


# Create land/country basemap from https://www.naturalearthdata.com/
land_sp <- ne_download(scale = 10, type = 'countries', category = 'cultural', returnclass = 'sp')
land_df <- fortify(land_sp) # covert to dataframe for ggplot


# What are the lat & long bound of closure poly
summary(closure_poly_sp)$bbox

### Map of Closure Area ###
Map_closure_trawl_longline <- ggplot() +
  geom_polygon(data = land_df, # basemap 
               aes(x = long, y = lat, group = group),
               fill = 'grey', color = "black", lwd = 0.1) +
  geom_point(data = fishing_trawl_long, # fishing datapoints
             aes(x = lon_bin, y = lat_bin, color = fishing_class),
             alpha = 0.5) + # make points transparent to show density if layered
  scale_colour_viridis_d(option = "viridis", # use viridis color scale for gear type
                         name = "Fishing Gear", 
                         labels = c("Drifting Longlines", "Set Longlines", "Trawlers")) +
  geom_polygon(data = closure_poly_df, # closure polygon
               aes(x = long, y = lat), color = 'black', fill = NA) +
  coord_sf(xlim = c(27, 37), ylim = c(30, 37), # add a few degrees to closure poly bounds
           crs = "+proj=longlat +datum=WGS84 +no_defs",
           expand = FALSE) +
  labs(x = "Longitude", 
       y = "Latitude",
       title = "Longline and Trawler Fishing 2019: Closure") +
  theme_classic(base_size = 12) +
  theme(legend.key = element_rect(fill = NA, color = NA), # remove fill color from legend
        panel.background = element_rect(fill = "aliceblue"))


Map_closure_trawl_longline

# Save map to Rproject
# ggsave(Map_closure_trawl_longline, 
#        filename = file.path(out_path, "Map_Closure_trawl_longline_2019.png"),
#        width = 10,
#        units = c("in"),
#        dpi = 400)
```

```{r Q2_other_closure, results='markup'}
# What are the lat & long bound of other closure poly
summary(other_poly_sp)$bbox

Map_other_trawl_longline <- ggplot() +
  geom_polygon(data = land_df, # land mass basemap
               aes(x = long, y = lat, group = group),
               fill = 'grey', color = "black", lwd = 0.1) +
  geom_point(data = fishing_trawl_long, # fishing datapoints
             aes(x = lon_bin, y = lat_bin, color =fishing_class),
             alpha = 0.5) + # make points transparent to show density if layered
  scale_colour_viridis_d(option = "viridis", # use viridis color scale for gear type
                         name = "Fishing Gear", 
                         labels = c("Drifting Longlines", "Set Longlines", "Trawlers")) +
  geom_polygon(data = other_poly_df, # other closure area poly
               aes(x = long, y = lat), 
               color = 'black', fill = NA) +
  coord_sf(xlim = c(3, 10), ylim = c(40, 45), 
           crs = "+proj=longlat +datum=WGS84 +no_defs",
           expand = FALSE) + # don't add extra to edges of limits
  labs(x = "Longitude", 
       y = "Latitude",
       title = "Longline and Trawler Fishing 2019: Other Closure") +
  theme_classic(base_size = 12) +
  theme(legend.key = element_rect(fill = NA, color = NA), # remove fill color from legend
        panel.background = element_rect(fill = "aliceblue"))

Map_other_trawl_longline

# Save map to Rproject
# ggsave(Map_other_trawl_longline, 
#        filename = file.path(out_path, "Map_Other_trawl_longline_2019.png"),
#        width = 10,
#        units = c("in"),
#        dpi = 400)
```

## Next Steps

1. Add inset maps
2. Polish output tables
3. Add tables to repo README